---
title: "4su_effect_1vs6_1vs2"
author: "Maddie Lombardo"
date: "2022-08-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

going to look at controls 1 vs 6hr and bdnf 1 vs 6hr
trying to normalize for the 4su effect 
```{r de controls 1vs6hr and bdnf 1vs6hr}
source(here::here('run_standard_deseq.R'))

feature_counts_fp <- file.path(here::here('data','feature_counts_bdnf_No4'))

one_vs_6_dds_control <- run_standard_deseq(feature_counts_fp, 
base_grep = "1hr",
                              contrast_grep = "6hr", 
                              grep_pattern = "CONTROL_.*.*_1|CONTROL_.*.*_6",
                              baseName = "onehour",
                              contrastName = '6hr')

one_vs_6_dds_bdnf <- run_standard_deseq(feature_counts_fp, 
base_grep = "1hr",
                              contrast_grep = "6hr", 
                              grep_pattern = "BDNF_.*.*_1|BDNF_.*.*_6",
                              baseName = "onehour",
                              contrastName = '6hr')
                          
#volcano plots
source(here::here('make_volcano_plot.R'))

label_significant(one_vs_6_dds_control$results_table, log2FoldCut = 1.5, log10padj = 4)+
  scale_fill_manual(values = c("#3fa7d6", "#ee6352", "#fac05e"), name = NULL)+
  scale_color_manual(values = c("#3fa7d6", "#ee6352", "#fac05e"), name = NULL, guide = FALSE)+
  ggtitle("The effect of 4SU in Control Condition", subtitle ="log2FoldChange(6 hr / 1 hr) vs log10p-value(6 hr / 1hr)")+
  ylab(expression(paste("-Lo", g[10], "P-value")))+
  xlab(expression(paste("Lo", g[2], "Fold Change")))+
  ggpubr::theme_pubr()

label_significant(one_vs_6_dds_bdnf$results_table, log2FoldCut =2, log10padj = 6)+
  scale_fill_manual(values = c("#3fa7d6", "#ee6352", "#fac05e"), name = NULL)+
  scale_color_manual(values = c("#3fa7d6", "#ee6352", "#fac05e"), name = NULL, guide = FALSE)+
  ggtitle("The effect of 4SU in BDNF Condition", subtitle ="log2FoldChange(6 hr / 1 hr) vs log10p-value(6 hr / 1hr)")+
  ylab(expression(paste("-Lo", g[10], "P-value")))+
  xlab(expression(paste("Lo", g[2], "Fold Change")))+
  ggpubr::theme_pubr()
```

1vs6 hr venn diagrams for control vs bdnf upreg/downreg
```{r 1vs6 venn diagram upreg/downreg}
library("ggVennDiagram")

upreg_genes_6hr_bdnf <- one_vs_6_dds_bdnf$results_table %>% 
  select(log2FoldChange,padj,gene_name) %>% 
  filter(log2FoldChange > 0.1 & padj < 0.1)

upreg_genes_6hr_control <- one_vs_6_dds_control$results_table %>% 
  select(log2FoldChange,padj,gene_name) %>% 
  filter(log2FoldChange > 0.1 & padj < 0.1)

upreg_gene_6hr_list <- list(A = upreg_genes_6hr_bdnf$gene_name,
                            B = upreg_genes_2hr_control$gene_name)

ggVennDiagram(upreg_gene_6hr_list, category.names = c("BDNF", "Control"))+
  ggtitle("Upregulated Genes 6hr BDNF vs Control")

downreg_genes_6hr_bdnf <-one_vs_6_dds_bdnf$results_table %>% 
  select(log2FoldChange,padj,gene_name) %>% 
  filter(log2FoldChange < -0.1 & padj < 0.1)

downreg_genes_6hr_control <- one_vs_6_dds_control$results_table %>% 
  select(log2FoldChange,padj,gene_name) %>% 
  filter(log2FoldChange < -0.1 & padj < 0.1)

downreg_gene_6hr_list <- list(A = downreg_genes_6hr_bdnf$gene_name,
                            B = downreg_genes_6hr_control$gene_name)

ggVennDiagram(downreg_gene_6hr_list, category.names = c("BDNF", "Control"))+
  ggtitle("Downregulated Genes 6hr BDNF vs Control")
```

now going to 1vs 2 hour for controls and bdnf
```{r 1vs2 hr controls and bdnf}
source(here::here('run_standard_deseq.R'))

feature_counts_fp <- file.path(here::here('data','feature_counts_bdnf_No4'))

one_vs_2_dds_control <- run_standard_deseq(feature_counts_fp, 
base_grep = "1hr",
                              contrast_grep = "2hr", 
                              grep_pattern = "CONTROL_.*.*_1|CONTROL_.*.*_2",
                              baseName = "onehour",
                              contrastName = '2hr')

one_vs_2_dds_bdnf <- run_standard_deseq(feature_counts_fp, 
base_grep = "1hr",
                              contrast_grep = "2hr", 
                              grep_pattern = "BDNF_.*.*_1|BDNF_.*.*_2",
                              baseName = "onehour",
                              contrastName = '2hr')
                          
#volcano plots
source(here::here('make_volcano_plot.R'))

label_significant(one_vs_2_dds_control$results_table, log2FoldCut = 1.5, log10padj = 4)+
  scale_fill_manual(values = c("#3fa7d6", "#ee6352", "#fac05e"))+
  scale_color_manual(values = c("#3fa7d6", "#ee6352", "#fac05e"))+
  ggtitle("Volcano 1 vs 2hr Control")+
  ggpubr::theme_pubr()

label_significant(one_vs_2_dds_bdnf$results_table, log2FoldCut =2, log10padj = 6)+
  scale_fill_manual(values = c("#3fa7d6", "#ee6352", "#fac05e"))+
  scale_color_manual(values = c("#3fa7d6", "#ee6352", "#fac05e"))+
  ggtitle("Volcano 1 vs 2hr BDNF")+
  ggpubr::theme_pubr()
```

gonna make venn diagrams of BDNF 2 hours v Control 2 hours (total RNA) up regulated, downregulated

```{r venn diagrams 2hr bdnf vs 2hr control}
library("ggVennDiagram")

upreg_genes_2hr_bdnf <- one_vs_2_dds_bdnf$results_table %>% 
  select(log2FoldChange,padj,gene_name) %>% 
  filter(log2FoldChange > 0 & padj < 0.1)

upreg_genes_2hr_control <- one_vs_2_dds_control$results_table %>% 
  select(log2FoldChange,padj,gene_name) %>% 
  filter(log2FoldChange > 0.1 & padj < 0.1)

upreg_gene_2hr_list <- list(A = upreg_genes_2hr_bdnf$gene_name,
                            B = upreg_genes_2hr_control$gene_name)

ggVennDiagram(upreg_gene_2hr_list, category.names = c("BDNF", "Control"))+
  ggtitle("Upregulated Genes 2hr BDNF vs Control")

downreg_genes_2hr_bdnf <-one_vs_2_dds_bdnf$results_table %>% 
  select(log2FoldChange,padj,gene_name) %>% 
  filter(log2FoldChange < -0.1 & padj < 0.1)

downreg_genes_2hr_control <- one_vs_2_dds_control$results_table %>% 
  select(log2FoldChange,padj,gene_name) %>% 
  filter(log2FoldChange < -0.1 & padj < 0.1)

downreg_gene_2hr_list <- list(A = downreg_genes_2hr_bdnf$gene_name,
                            B = downreg_genes_2hr_control$gene_name)

ggVennDiagram(downreg_gene_2hr_list, category.names = c("BDNF", "Control"))+
  ggtitle("Downregulated Genes 2hr BDNF vs Control")
```

looking at effect of 4su on rna expression using GO analysis

1vs6 control upregulated
```{r go on 1vs6 hr control: upregulated}
bg_ids <- one_vs_6_dds_control$results_table$Geneid %>% 
  gsub("\\..*", "", .) %>% 
  unique() %>% 
  clipr::write_clip()

upreg_1vs6_control <- one_vs_6_dds_control$results_table %>% 
  filter(log2FoldChange > 0.5 & padj < 0.1) %>% 
  pull(Geneid) %>% 
  gsub("\\..*", "", .) %>% 
  clipr::write_clip()

#graphing the string data
bp_go_1vs6_upreg_fp <- file.path("data", 'GO_analysis_from_string',"bp go 1vs6 control upreg.tsv")
bp_go_1vs6_upreg <- janitor::clean_names(read_tsv(bp_go_1vs6_upreg_fp))

bp_go_1vs6_upreg %>% slice_max(strength, n = 10) %>% 
  mutate(term_description = fct_reorder(term_description,strength)) %>% 
  ggplot(aes(y = strength, fill = false_discovery_rate, x = term_description,size = observed_gene_count))+
  geom_point(pch = 21) +
  scale_fill_gradient(low = 'blue', high = 'red')+
  coord_flip()+ 
  ggpubr::theme_pubr()+
  theme(legend.position = 'right')+
  ggtitle("biological process GO on 1vs6_upreg genes")

cc_go_1vs6_upreg_fp <- file.path("data",'GO_analysis_from_string', "cc go 1vs6 control upreg.tsv")
cc_go_1vs6_upreg <- janitor::clean_names(read_tsv(cc_go_1vs6_upreg_fp))

cc_go_1vs6_upreg %>% slice_max(strength, n = 10) %>% mutate(term_description = fct_reorder(term_description,strength)) %>%  ggplot(aes(y = strength, fill = false_discovery_rate, x = term_description,size = observed_gene_count)) + geom_point(pch = 21) + scale_fill_gradient(low = 'blue', high = 'red')+coord_flip()+ theme_pubr()+theme(legend.position = 'right')+
  ggtitle("Cellular Compartment GO on 1vs6_upreg genes")

mf_go_1vs6_upreg_fp <- file.path("data", "mf go 1vs6 control upreg.tsv")
mf_go_1vs6_upreg <- janitor::clean_names(read_tsv(mf_go_1vs6_upreg_fp))

mf_go_1vs6_upreg %>% slice_max(strength, n = 10) %>% mutate(term_description = fct_reorder(term_description,strength)) %>%  ggplot(aes(y = strength, fill = false_discovery_rate, x = term_description,size = observed_gene_count)) + geom_point(pch = 21) +scale_fill_gradient(low = 'blue', high = 'red')+ coord_flip()+ theme_pubr()+ theme(legend.position = 'right')+ggtitle("Molecular function GO on 1vs6_upreg genes")
```

1vs6 control downregulated
```{r go on 1vs6 hr control: downregulated}
downreg_1vs6_control <- one_vs_6_dds_control$results_table %>% 
  filter(log2FoldChange < -0.5 & padj < 0.1) %>% 
  pull(Geneid) %>% 
  gsub("\\..*", "", .) %>% 
  clipr::write_clip()

cc_go_1vs6_downreg_fp <- file.path("data", "cc go 1vs6 control downreg.tsv")
cc_go_1vs6_downreg <- janitor::clean_names(read_tsv(cc_go_1vs6_downreg_fp))

cc_go_1vs6_downreg %>% slice_max(strength, n = 10) %>% mutate(term_description = fct_reorder(term_description,strength)) %>%  ggplot(aes(y = strength, fill = false_discovery_rate, x = term_description,size = observed_gene_count)) + geom_point(pch = 21) + scale_fill_gradient(low = 'blue', high = 'red')+ coord_flip()+ theme_pubr()+ theme(legend.position = 'right')+
  ggtitle("Cellular Compartment GO on 1vs6 downreg genes")
```

```{r looking at dynein related genes}
dyn_cc_1vs6_upreg <- cc_go_1vs6_upreg %>% filter(grepl('dyn', term_description,ignore.case = TRUE)) %>% pull(matching_proteins_in_your_network_labels) %>% 
  str_split(.,',')

dyn_mf_1vs6_upreg <- mf_go_1vs6_upreg %>% filter(grepl('dyn',term_description,ignore.case = TRUE)) %>% 
  pull(matching_proteins_in_your_network_labels) %>% 
  str_split(.,',') 

dyn_1vs6_control <- unique(dyn_cc_1vs6_upreg, dyn_mf_1vs6_upreg[[1]] )[[1]]

normalized_counts <- one_vs_6_dds_control$deseq_obj %>% 
  counts(., normalized=TRUE) %>% 
  as.data.frame() %>% 
  rownames_to_column('ensgene') %>% 
  mutate(ensgene = gsub("\\..*", "",ensgene)) %>% 
  left_join(annotables::grch38 %>% select(ensgene,symbol)) %>% 
  filter(symbol %in% dyn_1vs6_control)

normalized_counts <- normalized_counts %>% 
  melt(id.vars = c('ensgene','symbol'))

normalized_counts <- normalized_counts %>% 
  mutate(condition = case_when(grepl('_1hr', variable, ignore.case = TRUE) ~ '1',
    grepl('_2hr', variable, ignore.case = TRUE) ~ '2',
    grepl('_6hr', variable, ignore.case = TRUE) ~ '6'))
   
ggplot(normalized_counts, mapping = aes(condition, value))+
  geom_boxplot()+
  geom_point()+
  geom_text_repel(mapping = aes(label = variable), size = 2.5)+
  facet_wrap(vars(symbol), scales = 'free_y')+
  ggpubr::theme_pubr()+
  labs(x = "Timepoint 1hr vs 6hr", y ="Normalized Counts" )
```

```{r scattterplot log2foldchange total rna}
up_4su_control_6hr <- one_vs_6_dds_control$results_table %>% filter(padj < 0.1 & log2FoldChange > 0)

test = list(up_4su_control_6hr = up_4su_control_6hr$gene_name,upreg_genes_2hr_bdnf = upreg_genes_2hr_bdnf$gene_name)
ggVennDiagram(test, category.names = c("Upregulated by 4SU (6 v 1hr Controls)","Upregulated by BDNF 2hr")) + ggtitle("Effect of 4SU treatment on total gene expression") +
    scale_color_manual(values = c("black","black"))

new_ratio_bayesian_fp <- file.path(here::here('data', 'new_ratio_bayesian_p_de.csv'))
new_ratio_bayesian <- read_csv(new_ratio_bayesian_fp)

new_rna_sig_bit <- new_ratio_bayesian %>% 
  select(gene, new_rna_sig)

names(new_rna_sig_bit)[names(new_rna_sig_bit) == 'gene'] <- 'Geneid'

new_rna_sig_bit<-new_rna_sig_bit %>% 
  filter(!is.na(new_rna_sig))

new_rna_sig_bit <- new_rna_sig_bit %>% 
  mutate(new_rna_sig = case_when(
    new_rna_sig == 'unclear' ~ "Unclear",
    new_rna_sig == 'bdnf_equals_control' ~ "BDNF equals Control",
    new_rna_sig == 'bdnf_higher_new_rna' ~ "BDNF higher new RNA",
    new_rna_sig == 'bdnf_lower_new_rna' ~ "BDNF lower new RNA"
  ))


one_vs_6_dds_control$results_table %>% 
    full_join(one_vs_2_dds_bdnf$results_table, by = c("Geneid"),
              suffix = c("_4su_effect","_BDNF_2hr")) %>% 
  left_join(new_rna_sig_bit, by = c('Geneid')) %>% 
  filter(!is.na(new_rna_sig)) %>% 
    ggplot(aes(x = log2FoldChange_BDNF_2hr,
               y = log2FoldChange_4su_effect, color = new_rna_sig)) + 
    geom_hline(yintercept = 0) +
    geom_vline(xintercept = 0) +
    geom_point(alpha = 0.4) + 
  facet_wrap(vars(new_rna_sig))+
  labs(color = 'New RNA')+
  ggpubr::theme_pubr() +
    ggtitle("Effect on Total RNA with 4SU Treatment (control 6 v 1 hr) vs BDNF 2hr",subtitle = 'All genes') + 
    ylab("Log2Fold Total RNA - 4SU Treatment") +
    xlab("Log2Fold Total RNA - BDNF 2hr")
```

going to add column:4su log2fold change/bdnf+4su log2foldchange to see if bdnf addition amplifies some of the seen 4su response
```{r bdnf+4su log2foldchange - 4su log2fold change}

one_vs_2_dds_bdnf_results <- one_vs_2_dds_bdnf$results_table %>% 
  full_join(one_vs_6_dds_control$results_table, by = c("Geneid"),
              suffix = c("_BDNF_2hr","_4su_effect_control"))

one_vs_2_dds_bdnf_results <- one_vs_2_dds_bdnf_results %>% 
  mutate(log2fold_ratio = log2FoldChange_4su_effect_control/log2FoldChange_BDNF_2hr) %>% 
  filter(!gene_name_BDNF_2hr == 'AC126474.1') %>% 
  filter(!(log2FoldChange_BDNF_2hr < 0 & log2FoldChange_4su_effect_control > 0)) %>% 
  filter(!(log2FoldChange_BDNF_2hr > 0 & log2FoldChange_4su_effect_control < 0)) %>% 
  mutate(log2fold_ratio = ifelse(log2FoldChange_BDNF_2hr<0, log2fold_ratio*-1, log2fold_ratio))

ggplot(one_vs_2_dds_bdnf_results, aes(log2fold_ratio))+
         geom_density()+
  xlim(-25,25)+
  ggpubr::theme_pubr()+
  ggtitle("Relative gene expression effect of 4su vs BDNF in total RNA")+
  xlab("log2FoldChange_4su / log2FoldChange_BDNF")

background_genes_overlap <- unique(c(one_vs_6_dds_control$results_table$Geneid,one_vs_2_dds_bdnf$results_table$Geneid)) %>% 
    gsub("\\..*", "", .) %>% 
  clipr::write_clip()

one_vs_2_dds_bdnf_results %>% 
  arrange(-log2fold_difference) %>% 
  slice_max(order_by = log2fold_difference, n = 2000) %>% 
  pull(Geneid) %>% 
  gsub("\\..*", "", .) %>% 
  clipr::write_clip()

BP_GO_2hr_bdnf_minus_4su_fp <- file.path("data", "2bdnf-4su_GO_BP.tsv")
BP_GO_2hr_bdnf_minus_4su <- janitor::clean_names(read_tsv(BP_GO_2hr_bdnf_minus_4su_fp))

BP_GO_2hr_bdnf_minus_4su %>% slice_max(strength, n = 10) %>% 
  mutate(term_description = fct_reorder(term_description,strength)) %>% 
  ggplot(aes(y = strength, fill = false_discovery_rate, x = term_description,size = observed_gene_count))+
  geom_point(pch = 21) +
  scale_fill_gradient(low = 'blue', high = 'red')+
  coord_flip()+ 
  theme_pubr()+
  theme(legend.position = 'right')+
  ggtitle("The effect of BDNF accounting for 4su changes", subtitle = "BP GO terms for BDNF +4su at 2hr minus 4su at 6hr")

CC_GO_2hr_bdnf_minus_4su_fp <- file.path("data", "2bdnf-4su_GO_CC.tsv")
CC_GO_2hr_bdnf_minus_4su <- janitor::clean_names(read_tsv(CC_GO_2hr_bdnf_minus_4su_fp))

CC_GO_2hr_bdnf_minus_4su %>% slice_max(strength, n = 10) %>% mutate(term_description = fct_reorder(term_description,strength)) %>%  ggplot(aes(y = strength, fill = false_discovery_rate, x = term_description,size = observed_gene_count)) + geom_point(pch = 21) + scale_fill_gradient(low = 'blue', high = 'red')+coord_flip()+ theme_pubr()+theme(legend.position = 'right')+
  ggtitle("The effect of BDNF accounting for 4su changes", subtitle = "BP GO terms for BDNF +4su at 2hr minus 4su at 6hr")

mf_go_1vs6_upreg_fp <- file.path("data", "mf go 1vs6 control upreg.tsv")
mf_go_1vs6_upreg <- janitor::clean_names(read_tsv(mf_go_1vs6_upreg_fp))

mf_go_1vs6_upreg %>% slice_max(strength, n = 10) %>% mutate(term_description = fct_reorder(term_description,strength)) %>%  ggplot(aes(y = strength, fill = false_discovery_rate, x = term_description,size = observed_gene_count)) + geom_point(pch = 21) +scale_fill_gradient(low = 'blue', high = 'red')+ coord_flip()+ theme_pubr()+ theme(legend.position = 'right')+ggtitle("Molecular function GO on 1vs6_upreg genes")

```
