```{r setup}
knitr::opts_chunk$set(echo = TRUE)
library("tidyverse")
library("DESeq2")
library("ggplot2")
library("pacman")
library("readr")
library('data.table')
```

this let's you access the 'create_feature_count_table' function gifted by AL :)

```{r}
source(here::here('create_feature_count_table.R'))
```

this organizes all the data from the many .bam files and creates one data frame you can work with

```{r}
featureCounts <- create_feature_count_table("C:/Users/mlomb/OneDrive/Desktop/MRes project/Differential Expression/data/bdnf_trial_full/bdnf_trial_full")
```

this lets you access the 'make_deseq_dfs.R'

```{r}
source(here::here('make_deseq_dfs.R'))
```

this runs the make_deseq_dfs() function

```{r}
my_df <- make_deseq_dfs(featureCounts, grep_pattern = "", base_grep = "CONTROL", contrast_grep = "BDNF")
```

accessing 'run_standard_deseq.R'

```{r}
source(here::here('run_standard_deseq.R'))
```

running run_standard_deseq

```{r}
my.dds <- run_standard_deseq("C:/Users/mlomb/OneDrive/Desktop/MRes project/Differential Expression/data/bdnf_trial_full/bdnf_trial_full", 
                               base_grep = "CONTROL",
                               contrast_grep = "BDNF", 
                              grep_pattern = "",
                              baseName = "control",
                              contrastName = 'BDNF')
```

accessing volcano plot making function

```{r}
source(here::here('make_volcano_plot.R'))
```

running it to make preeeettttyyyy graph

```{r}
my_plot <- make_volcano_plot(my.dds$results_table)
```
