---
title: "create deseq table from grandslam data"
author: "Maddie Lombardo"
date: "2/15/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(PCAtools)
library(gprofiler2)
library(pcaExplorer)
library(clusterProfiler)
library(RColorBrewer)
library(ggpubr)
library(GOfuncR)
```

This function will be organizing this data into 2 tables that can be input into a Deseq
I will be creating a table 1 that will act as the counts data frame that includes the gene and the min2 counts
Table 2 will act as the metadata for the deseq

```{r reading in file path}
grandslam_new_RNA_file_path <- file.path(here::here('data/grandslam_new_rna.csv'))

grandslam_new_rna <- read_csv(grandslam_new_RNA_file_path)
```

```{r filtering out any estimate new credible level > 20%}
grandslam_new_rna_filtered <- grandslam_new_rna %>% 
  filter(estimate_new_credible_interval < 0.2)
```

creating table 1
```{r filter conversions at specified timepoint and make table 1 from it}
create_table_1 <- function(filtered_grandslam_rna, 
                          timepoint = "6"){
    
  filtered_df <- filtered_grandslam_rna %>% 
      filter(time == timepoint) %>% 
      select(gene, min2, SampleID) %>% 
      mutate(min2 = round(min2)) %>% 
      pivot_wider(names_from = SampleID, values_from = min2, values_fill = 0) 
    
    return(filtered_df)
  }
                         
```

creating table 2
```{r make table 2 for metadata table}
create_metatable <- function(table_1){
  
  meta_table <- tibble(sample = table_1 %>%  
                         colnames()) %>% 
    separate(sample, into = c("condition", "well", "time"), remove = FALSE) %>%
    column_to_rownames('sample') %>% 
    mutate(factor_name = factor(condition, levels = c("control","bdnf")))
  
  return(meta_table)
}
```

now for the deseq part
```{r deseq}
library("DESeq2")

min2_deseq <- function(my_table, time){
  
  table_1 <- create_table_1(my_table,time)
  
  table_1 <- table_1 %>% column_to_rownames('gene')
  
  meta_table <- create_metatable(table_1)
  
  dds_min2 <- DESeqDataSetFromMatrix(countData = table_1,
                                   colData = meta_table,
                                   design = ~ factor_name)
  dds_min2 <- DESeq(dds_min2)

  


  res_min2 <- results(dds_min2) %>% 
    as.data.frame() %>% 
    rownames_to_column('ensgene') %>% 
    mutate(ensgene = gsub("\\..*", "", ensgene)) %>%
    left_join(annotables::grch38 %>% select(ensgene,symbol))
    
    



return(res_min2)
}
```

testing function on time 2
```{r test new function on time 2}
time2 <- min2_deseq(grandslam_new_rna_filtered, time = 2)

time2 %>% 
  ggplot(mapping = aes(x = pvalue)) + 
  geom_histogram()
```

using function on time 1 (it works!)
```{r time 1 pvalue chart}
time1 <- min2_deseq(grandslam_new_rna_filtered, time =1)

time1 %>% 
  ggplot(mapping = aes(x = pvalue)) + 
  geom_histogram()
```

time 6
```{r time 6 pvalue chart}

time6 <- min2_deseq(grandslam_new_rna_filtered, time =6)

time6 %>% 
  ggplot(mapping = aes(x = pvalue)) + 
  geom_histogram()
```

make 1 table
```{r squish together} 
de_list <- list(time1,time2,time6) #make a list of whatever you called the 3 results
samp_ids <- c(1,2,6)  #one two six
full_de_min2 <- purrr::map2(de_list, samp_ids, ~cbind(.x, time = .y)) %>% 
    data.table::rbindlist(.)
```

```{r read in bayesian shizz}
bayesian_fp <- file.path("data","new_ratio_bayesian_p_de.csv")
new_ratio_p <- read_csv(bayesian_fp)
```


we making a log2foldchange graph comparing new to total
```{r log2fold new vs total}
significant_genes <- full_de_min2 %>%
  filter(log2FoldChange_total_rna > 2 & padj_total_rna < 0.1) %>% 
  pull(gene_name_total_rna)

full_de_min2 <- full_de_min2 %>% 
  left_join(new_ratio_p %>% 
              mutate(ensgene = gsub("\\..*", "", gene)), 
            by = c('ensgene','time'),
            suffix = c("_total_rna","_new_rna")) %>%
  filter(!is.na(padj_total_rna)) %>% 
 mutate(significant_log2fold_total = padj_total_rna < 0.1) 

my_color_palette = c("#4cbab3", "#ee6352", "#59cd90", "#fac05e", "#3fa7d6")
  
full_de_min2 %>% 
  mutate(plot_name = ifelse(log2FoldChange_total_rna > 3 & padj_total_rna < 0.05, gene_name_total_rna, NA)) %>% 
  ggplot(aes(x = log2FoldChange_total_rna,
             y = log2FoldChange_new_rna,
             fill =significant_log2fold_total, 
             color = significant_log2fold_total))+
  geom_hline(yintercept = 0)+
  geom_vline(xintercept = 0)+
  geom_point(pch = 21)+
  scale_colour_manual(values = my_color_palette)+
  geom_jitter()+
  facet_wrap(vars(time))+
  geom_text_repel(aes(label = plot_name), color = "black")

##4 - Compare results here to the ```{r categorizing new vs total} in "differential expression on coverted" analysis
```
explaining the graph quadrants:
close to center: these genes do not have large/significant changes in expression
1(top left): these genes have a log 2fold change increase in new rna but a decrease in total rna
2(top right): these genes have both a log2fold change increase in new and total rna
3(bottom right): these genes have an increase change in total rna but a decreased change in new fraction rna
4(bottom left): these genes have both a decrease change in total and new rna

comparing to categorizing new vs total:
Are there any exciting new genes or categories of genes? 