---
title: "how to DESeq2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("tidyverse")
library("DESeq2")
library("ggplot2")
library("pacman")
library("readr")
library('data.table')
```

First, load in the counts file you are using:
-update the file name you are using

```{r file upload}
counts_file <- here::here("data", "4su_baseline.xlsx")
cts <- readxl::read_xlsx(counts_file)
```

Next, set up your counts and metadata for DESeq 

```{r cts, meta_df }
cts <- cts %>% column_to_rownames('ensgene')
meta_df <- data.frame(samples = colnames(cts))
meta_df <- meta_df %>% 
  filter(samples != "ensgene") %>%
  mutate(condition = ifelse(grepl("F",samples), "f210i_mut", "control")) %>% 
  column_to_rownames('samples')

cts <- cts[, rownames(meta_df)]
all(rownames(meta_df)==colnames(cts))
```

Yay, now it's time to run the DESeq
```{r DESeq2}
dds <- DESeqDataSetFromMatrix(countData = cts,
                              colData = meta_df,
                              design = ~ condition)

keep <- rowSums(counts(dds)) >= 10
dds <- dds[keep,]

dds <- DESeq(dds)
res <- results(dds)
```

If you want to order the results of the DESeq by the pvalue,

```{r Order res}
resOrdered <- res[order(res$pvalue),]
summary(res)
sum(res$padj < 0.1, na.rm = TRUE)
```

You can adjust the alpha level used (above uses the standard, 0.1)

```{r adjust alpha}
res05 <- results(dds, alpha = 0.05)
summary(res05)
sum(res$padj < 0.05, na.rm = TRUE)
```

Now to make some plots--wooooo!
This is how to make an MA plot. You will notice that you shrink the data before, this is because it is more useful to visualize the shrunken log2 fold changes because it removed the noise associated w/ log2 fold changes from low count genes w/o requiring arbitrary filtering thresholds

```{r make MA plot}
resLFC <- lfcShrink(dds, coef = "condition_f210i_mut_vs_control", type = "apeglm")
plotMA(resLFC, ylim=c(-2,2))
```

To examine the counts of reads for the most significant gene across the groups in a plot,

```{r plot counts}
d <- plotCounts(dds, gene=which.min(res$padj), intgroup = "condition", returnData = TRUE)

my_ggplot <- ggplot(d, aes(x = condition, y = count)) +
  geom_point(position = position_jitter(w=0,h=0))+
  scale_y_log10(breaks=c(25,100,400))

print(my_ggplot)
```

Now to make this actually useful to you, we are going to grab the gene name and put it as the title to the plot.
**NOTE: make sure you are using the right annotables, this one uses the mouse (grcm38). human is (grch38)

```{r grab symbol and make title}
d %>% 
  rownames_to_column("ensgene") %>% 
  as.data.frame() %>% 
  left_join(annotables::grcm38) %>% 
  select(symbol, count, condition, ensgene) %>% 
  melt(id.vars= c("symbol", "condition"))
  
title <- resLFC %>% 
  as.data.frame() %>% 
  rownames_to_column("ensgene") %>% 
  left_join(annotables::grcm38) %>% 
  filter(padj == min(resLFC$padj, na.rm = TRUE)) %>% 
  pull(symbol)

ggplot(d, aes(x= condition, y=count, fill = condition))+
  geom_point(position = position_jitter(w=0,h=0))+
  scale_y_log10(breaks=c(25,100,400)) +
  ggtitle(title)
```

After all of this you can make Volcano plots (ooooooo) using the make_volcano_plot.R function