---
title: "PCA on the 4SU Trial Data"
author: "Anna-Leigh Brown"
date: "18/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
if (!require("pacman")){
    install.packages("pacman")
} #if pacman is not installed, install it
pacman::p_load(tidyverse, data.table, ggplot2,janitor,ggpubr,ggrepel,
               devtools,readxl) #pacman will now install and load these packages if you don't have them

if (!require("annotables")) {
    devtools::install_github("stephenturner/annotables")
}
```

The chunk above is code. But this is just a normal text document. 

Let's load in the file:

I'm using a function called fread from the data.table package to read in the data. 
I do this cause it reads data faster, it's not relevant for a smaller table like
this one is, but can be handy once we get around to larger datasets. 

```{r}
bdnf_file_path <- file.path(here::here(),"data","bdnf_full_counts.csv")

bdnf_counts <-  read_csv(bdnf_file_path)

##bdnf_counts <- bdnf_counts %>% filter(ensgene !=  "ENSG00000251562")
```


As before we're going to put the "ensgene" into a 'rowname' and then 
do the PCA analysis.

Now I'm going to write this using the pipe operator "%>%". 

Your "hot key" to type that in is: Ctrl + Shift + M 

This command:

```{r}
pca_piped <- bdnf_counts %>% 
    column_to_rownames('ensgene') %>% 
    t() %>% 
    prcomp(.,scale = TRUE)
```


So let's do a PCA plot using meta data. You'll find formatted meta data in the drive

```{r}
path_to_meta_you_downloaded <-  file.path(here::here(),"data","trial_meta.csv")
meta_df <- fread(path_to_meta_you_downloaded)
meta_df <- meta_df %>% 
    column_to_rownames('sample') 

```

Let's read in the counts table and make a named version
```{r}
bdnf_counts_named <- bdnf_counts %>% 
    left_join(annotables::grch38 %>% dplyr::select(ensgene,symbol))

bdnf_counts <- bdnf_counts %>% 
    column_to_rownames('ensgene')
```


```{r}

library(PCAtools)
p <- PCAtools::pca(bdnf_counts, metadata = meta_df, removeVar = 0.1)
```
What does the removeVar argument do? Why did I include it? 

As before the screeplot
```{r}
screeplot(p, axisLabSize = 18, titleLabSize = 22)
```

Now we're going to make what is called an "eigencorplot"

This is going to correlate our metadata variables with our principle components. 
Star means "significant". What kind of correlation is it? Check out the help documentation 
by entering `?PCAtools::eigencorplot` in the console
```{r}
PCAtools::eigencorplot(p,components = getComponents(p, seq_len(6)),
                       metavars = c('labelling_time',"cond"))
```

Ohh my, PC1 (our most important variable) really is the labelling time. 

We can do a biplot with the 2 PC's that are correlating - e.g. labelling and 
PC5 which is the effect of BDNF

```{r}
biplot(p,
       x = "PC1",
       y = "PC5",
       colby = 'cond')
```

BDNF's effect seems to be really tightly on the first 2 hours of BDNF + 4SU. 

Let's go ahead and find out what the genes are that are correlated with PC1 - eg
what's happening at 6 hours labelling time? 

Let's directly take the loadings of each of the genes (a PCA is a linear combination). 

I'm going to also show you now how to put the symbol back on the table. 
We shall use the wonder of the `left_join` 

```{r}
pc1_gene_loadings = p$loadings %>% 
    dplyr::select(PC1) %>% 
    rownames_to_column('ensgene') %>% 
    left_join(annotables::grch38 %>% dplyr::select(ensgene,symbol)) %>% 
#https://statisticsglobe.com/r-dplyr-join-inner-left-right-full-semi-anti
    arrange(-PC1)
```



I want to look at the expression of the top 15 genes on PC1 so I'm going 
to plot their expression. 

Practice removing the pipes and see what every line does
```{r}
bdnf_counts %>% 
    rownames_to_column('ensgene') %>% 
    left_join(pc1_gene_loadings %>% head(15)) %>% 
    filter(!is.na(PC1)) %>% 
    select(-ensgene) %>% 
    melt(id.vars = c("symbol","PC1")) %>% 
    mutate(symbol = fct_reorder(symbol,-PC1)) %>% 
    separate(variable, into = c("condition","labelling_time")) %>% 
    ggplot(aes(x = labelling_time, y = value, fill = condition)) + 
    geom_col(position = "dodge2") + 
    facet_wrap(~symbol)

```



MAKE a barplot of the top 100 genes contributions and use that to choose cutoff

```{r}
top_100 <- pc1_gene_loadings %>% 
    slice_max(PC1, n = 100)

top_100 %>% 
     mutate(symbol = fct_reorder(symbol, PC1)) %>% 
     ggplot(aes(x = symbol, y = PC1)) + 
     geom_col(position = "dodge2") +
     coord_flip()
```

Given the above graph, I will take the top 30
TAKE THE TOP  GENES ON THE 4SU  PC1 and do GO analysis

```{r}
library(clusterProfiler)
library(org.Hs.eg.db)

go_4su <- pc1_gene_loadings %>% 
    slice_max(PC1, n=30)

ego <- enrichGO(gene = go_4su$ensgene,
                OrgDb = org.Hs.eg.db,
                ont = "ALL",
                keyType = "ENSEMBL",
                pAdjustMethod = "BH",
                pvalueCutoff = 0.01,
                qvalueCutoff = 0.05,
                readable = TRUE)

(go_dot <- dotplot(ego, showCategory = 30) +
    ggtitle("dotplot for PC1"))

```

TAKE THE TOP GENES ON THE BDNF PC5 and do GO analysis

```{r}
go_bdnf <- pc5
```

4su labelling is overwhelming the pc1 so i do analysis without those 2 timepoints, removing BDNF_6h, CONTROL_6h

```{r}
bdnf_counts <- select(bdnf_counts, -c(BDNF_6h, CONTROL_6h))

meta_df <- meta_df %>% filter(labelling_time != 6)

bdnf_counts <- select(bdnf_counts, -(symbol)) %>% 
    unique() %>% 
    remove_rownames() %>% 
    column_to_rownames("ensgene")
```

what genes account for PC3 w/o BDNF6h, Control6h

```{r}
pc3_gene_loadings = p$loadings %>% 
    dplyr::select(PC3) %>% 
    rownames_to_column('ensgene') %>% 
    left_join(annotables::grch38 %>% dplyr::select(ensgene,symbol)) %>% 
#https://statisticsglobe.com/r-dplyr-join-inner-left-right-full-semi-anti
    arrange(-PC3)

bdnf_counts %>% 
    rownames_to_column("ensgene") %>% 
    left_join(pc3_gene_loadings %>% head(15))  %>% 
    filter(!is.na(PC3)) %>% 
    select(-ensgene) %>% 
    melt(id.vars = c("symbol","PC3")) %>% 
    mutate(symbol = fct_reorder(symbol,-PC3)) %>% 
    separate(variable, into = c("condition","BDNF")) %>%
    ggplot(aes(x = BDNF, y = value, fill = condition)) +
    geom_col(position = "dodge2") + 
    facet_wrap(~symbol)
```



