---
title: "Differential expression no well 4"
author: "Maddie Lombardo"
date: "1/14/2022"
output:
  html_document: 
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(annotables)
library(PCAtools)
library(gprofiler2)
library(pcaExplorer)
library(AnnotationDbi)
library(org.Hs.eg.db)
library(clusterProfiler)
```

## Steps for Differential Expression 

### Reading the functions in that I'll be using in this document
```{r reading the functions in}
source(here::here('run_standard_deseq.R'))
source(here::here('make_volcano_plot.R'))
source(here::here('filter_significant.R'))
source(here::here('cluster_profiler_GO_analysis.R'))

```

### reading in the file
```{r read in files}
feature_counts_file_path <- file.path(here::here("data","feature_counts_bdnf_No4"))
```

## Creating the objects for GO analysis

### DeSeq 1 hr
```{r deseq2 1h}
one_hour_dds <- run_standard_deseq(feature_counts_file_path, base_grep = "CONTROL",
                              contrast_grep = "BDNF", 
                              grep_pattern = "1hr",
                              baseName = "control",
                              contrastName = 'BDNF')
```

### Make a volcano plot 1hr
```{r volcano plot 1h}
one_hour_volcano <- label_significant(one_hour_dds$results_table, log2FoldCut = 2.5, log10padj = 30)+
  ggtitle("Volcano Plot 1 Hour - Well 4 removed")

print(one_hour_volcano)
```

### pvalue chart for 1hr
```{r pvalue chart 1hr}
one_hour_dds$results_table %>% 
  ggplot(mapping = aes(x = pvalue)) + 
  geom_histogram()+
  ggtitle("P Value Chart - time 1hr - total rna")
```


### DeSeq 2hr
```{r deseq 2h}
two_hour_dds <- run_standard_deseq(feature_counts_file_path, base_grep = "CONTROL",
                              contrast_grep = "BDNF", 
                              grep_pattern = "2hr",
                              baseName = "control",
                              contrastName = 'BDNF')
```

### Make volcano plot 2hr
```{r volcano 2h}
two_hr_volcano <- label_significant(two_hour_dds$results_table, log2FoldCut = 2.5, log10padj = 30)+
  ggtitle("Volcano Plot 2 Hours - well 4 removed")

print(two_hr_volcano)
```

### P value chart 2hr
```{r}
two_hour_dds$results_table %>% 
  ggplot(mapping = aes(x = pvalue)) + 
  geom_histogram()+
  ggtitle("P Value Chart - time 2hr - total rna")
```

### DeSeq 6h
```{r deseq 6h}
six_hr_dds <- run_standard_deseq(feature_counts_file_path, base_grep = "CONTROL",
                              contrast_grep = "BDNF", 
                              grep_pattern = "6hr",
                              baseName = "control",
                              contrastName = 'BDNF')
```

### Make volcano plot 6h
```{r volcano 6h}
six_hr_volcano <- label_significant(six_hr_dds$results_table, log2FoldCut = 2.5, log10padj = 30)+
  ggtitle("Volcano Plot 6 Hours - well 4 removed")

print(six_hr_volcano)
```

### Make pvalue plot 6hr
```{r pvalue 6hr}
six_hr_dds$results_table %>% 
  ggplot(mapping = aes(x = pvalue)) + 
  geom_histogram()+
  ggtitle("P Value Chart time 6hr total rna")
```


## Gene Ontology Setup!

Now we are going to use the deseq data to do gene ontology at each of our time points (1,2, and 6 hours) for both upregulated and downregulated genes.

### GO 1 hr setup

This is grabbing both the up- and down- regulated genes for the 1 hour timepoint.
```{r GO 1hr}
results_1hr <- one_hour_dds$results_table

up_1hr_ensembl <- filter_significant(results_1hr,
                              padj_cutoff = 0.1,
                              log2Fold_cutoff = 0,
                              direction = "up")

down_1hr_ensembl <- filter_significant(results_1hr,
                              padj_cutoff = 0.1,
                              log2Fold_cutoff = 0,
                              direction = "down")
```

### Gostplot for upregulated genes at 1 hour
```{r up gostplot 1 hr}
gp_up_1hr = gost(up_1hr_ensembl, organism = "hsapiens", ordered_query = TRUE) 
gostplot(gp_up_1hr)
```

### Gostplot for downregulated genes at 1 hour
```{r down gostplot 1 hr}
gp_down_1hr = gost(down_1hr_ensembl, organism = "hsapiens", ordered_query = TRUE)
gostplot(gp_down_1hr)
```

### GO hr 2 Setup

This is grabbing both the up- and down- regulated genes for the 2 hour timepoint.
```{r go 2h}
results_2hr <- two_hour_dds$results_table

up_2hr_ensembl <- filter_significant(results_2hr,
                              padj_cutoff = 0.1,
                              log2Fold_cutoff = 0,
                              direction = "up")

down_2hr_ensembl <- filter_significant(results_2hr,
                              padj_cutoff = 0.1,
                              log2Fold_cutoff = 0,
                              direction = "down")
```

### Gostplot for upregulated genes at 2 hours
```{r up gostplot 2 hr}
gp_up_2hr <-  gost(up_2hr_ensembl, organism = "hsapiens")
gostplot(gp_up_2hr)
```

### Gostplot for downregulated genes at 2 hours
```{r down gostplot 2hr}
gp_down_2hr <-  gost(down_2hr_ensembl, organism = "hsapiens")
gostplot(gp_down_2hr)
```

### GO 6hr Setup

This is grabbing the up- and down- regulated genes at the 6 hour timepoint.
```{r go 6h}
results_6hr <- six_hr_dds$results_table

up_6hr_ensembl <- filter_significant(results_6hr,
                              padj_cutoff = 0.1,
                              log2Fold_cutoff = 0,
                              direction = "up")

down_6hr_ensembl <- filter_significant(results_6hr,
                              padj_cutoff = 0.1,
                              log2Fold_cutoff = 0,
                              direction = "down")
```

### Gostplot for upregulated genes at 6 hrs

```{r up gostplot 6hr}
gp_up_6hr = gost(up_6hr_ensembl, organism = "hsapiens")
gostplot(gp_up_6hr)
```

### Gostplot for downregulated genes at 6 hrs

```{r down gostplot 6hr}
gp_down_6hr = gost(down_6hr_ensembl, organism = "hsapiens")
gostplot(gp_down_6hr)
```

# its GO time!!!!

## GO 1 hr

### GO 1 hr Setup
```{r clusterprofiler GO 1hr, up and down}

e_go_cc_1hr_up <- cp_GO_analysis(up_1hr_ensembl, one_hour_dds, typeGO = "cc")

e_go_mf_1hr_up <- cp_GO_analysis(up_1hr_ensembl, one_hour_dds, typeGO = "mf")

e_go_bp_1hr_up <- cp_GO_analysis(up_1hr_ensembl, one_hour_dds, typeGO = "bp")

e_go_cc_1hr_down <- cp_GO_analysis(down_1hr_ensembl, one_hour_dds, typeGO = "cc")

e_go_mf_1hr_down <- cp_GO_analysis(down_1hr_ensembl, one_hour_dds, typeGO = "mf")

e_go_bp_1hr_down <- cp_GO_analysis(down_1hr_ensembl, one_hour_dds, typeGO = "bp")
```

### Cnet plot CC upregulated 1 hour
```{r cnet plot cc upregulated 1hr}
if(any(e_go_cc_1hr_up@result$p.adjust <0.01)){
  cc_cnet_1hr_up <- cnetplot(e_go_cc_1hr_up)+
    ggtitle("C Net Plot CC Upregulated 1 hr")
    print(cc_cnet_1hr_up)}
```

### dotplot CC upregulated 1hr

```{r CC dotplot upregulated 1hr}
if(any(e_go_cc_1hr_up@result$p.adjust <0.01)){
  cc_dotplot_1hr_up <- dotplot(e_go_cc_1hr_up)+
    ggtitle("Dotplot CC upregulated 1 hr")
print(cc_dotplot_1hr_up)}
```

### cnet plot BP upregulated 1hr
```{r BP cnet upregulated 1hr}
if(any(e_go_bp_1hr_up@result$p.adjust <0.01)){
  bp_cnetplot_1hr_up <- cnetplot(e_go_bp_1hr_up)+
    ggtitle("CNet Plot BP Upregulated 1hr")
print(bp_cnetplot_1hr_up)}
```

### dotplot BP upregulated 1hr

```{r BP dotplot upregulated 1hr}
if(any(e_go_bp_1hr_up@result$p.adjust <0.01)){
  bp_dotplot_1hr_up <- dotplot(e_go_bp_1hr_up)+
    ggtitle("Dot Plot BP Upregulated 1hr")
print(bp_dotplot_1hr_up)}
```

### cnet plot MF 1hr Upregulated

```{r mf cnet upregulated 1hr}
if(any(e_go_mf_1hr_up@result$p.adjust <0.01)){
  mf_cnetplot_1hr_up <- cnetplot(e_go_mf_1hr_up)+
    ggtitle("CNet Plot MF Upregulated 1hr")
print(mf_cnetplot_1hr_up)}
```

### dotplot MF 1hr Upregulated

```{r mf dotplot upregulated 1hr}
if(any(e_go_mf_1hr_up@result$p.adjust <0.01)){
  mf_dotplot_1hr_up <- dotplot(e_go_mf_1hr_up)+
    ggtitle("Dot Plot MF Upregulated 1hr")
print(mf_dotplot_1hr_up)}
```

### Cnet plot CC downregulated 1 hour
```{r cnet plot cc downregulated 1hr}
if(any(e_go_cc_1hr_down@result$p.adjust <0.01)){
  cc_cnet_1hr_down <- cnetplot(e_go_cc_1hr_down)+
    ggtitle("C Net Plot CC Down Regulated 1 hr")
  print(cc_cnet_1hr_down)}
```

### dotplot CC downregulated 1hr

```{r CC dotplot downregulated 1hr}
if(any(e_go_cc_1hr_down@result$p.adjust <0.01)){
  cc_dotplot_1hr_down <- dotplot(e_go_cc_1hr_down)+
    ggtitle("Dotplot CC Down regulated 1 hr")
  print(cc_dotplot_1hr_down)}
```

### cnet plot BP down regulated 1hr
```{r BP cnet down regulated 1hr}
if(any(e_go_bp_1hr_down@result$p.adjust <0.01)){
  bp_cnetplot_1hr_down <- cnetplot(e_go_bp_1hr_down)+
    ggtitle("CNet Plot BP down regulated 1hr")
  print(bp_cnetplot_1hr_down)}
```

### dotplot BP down regulated 1hr

```{r BP dotplot down regulated 1hr}
if(any(e_go_bp_1hr_down@result$p.adjust <0.01)){
  bp_dotplot_1hr_down <- dotplot(e_go_bp_1hr_down)+
    ggtitle("Dot Plot BP down regulated 1hr")
  print(bp_dotplot_1hr_down)}
```

### cnet plot MF 1hr down regulated

```{r mf cnet down regulated 1hr}
if(any(e_go_mf_1hr_down@result$p.adjust <0.01)){
  mf_cnetplot_1hr_down <- cnetplot(e_go_mf_1hr_down)+
    ggtitle("CNet Plot MF down regulated 1hr")
  print(mf_cnetplot_1hr_down)}
```

### dotplot MF 1hr down regulated

```{r mf dotplot down regulated 1hr}
if(any(e_go_mf_1hr_down@result$p.adjust <0.01)){
  mf_dotplot_1hr_down <- dotplot(e_go_mf_1hr_down)+
    ggtitle("Dot Plot MF down regulated 1hr")
print(mf_dotplot_1hr_down)}
```

## GO 2 hr

### GO 2 hr setup
```{r clusterProfiler GO at 2hr, up and down}
e_go_cc_2hr_up <- cp_GO_analysis(up_2hr_ensembl, two_hour_dds, typeGO = "cc")

e_go_mf_2hr_up <- cp_GO_analysis(up_2hr_ensembl, two_hour_dds, typeGO = "mf")

e_go_bp_2hr_up <- cp_GO_analysis(up_2hr_ensembl, two_hour_dds, typeGO = "bp")

e_go_cc_2hr_down <- cp_GO_analysis(down_2hr_ensembl, two_hour_dds, typeGO = "cc")

e_go_mf_2hr_down <- cp_GO_analysis(down_2hr_ensembl, two_hour_dds, typeGO = "mf")

e_go_bp_2hr_down <- cp_GO_analysis(down_2hr_ensembl, two_hour_dds, typeGO = "bp")
```


### Cnet plot CC upregulated 2 hour
```{r cnet plot cc upregulated 2hr}
if(any(e_go_cc_2hr_up@result$p.adjust <0.01)){
  cc_cnet_2hr_up <- cnetplot(e_go_cc_2hr_up)+
    ggtitle("C Net Plot CC Upregulated 2 hr")
  print(cc_cnet_2hr_up)}
```

### dotplot CC upregulated 2hr

```{r CC dotplot upregulated 2hr}
if(any(e_go_cc_2hr_up@result$p.adjust <0.01)){
  cc_dotplot_2hr_up <- dotplot(e_go_cc_2hr_up)+
    ggtitle("Dotplot CC upregulated 2 hr")
  print(cc_dotplot_2hr_up)}
```

### cnet plot BP upregulated 2hr
```{r BP cnet upregulated 2hr}
if(any(e_go_bp_2hr_up@result$p.adjust <0.01)){
  bp_cnetplot_2hr_up <- cnetplot(e_go_bp_2hr_up)+
    ggtitle("CNet Plot BP Upregulated 2hr")
  print(bp_cnetplot_2hr_up)}
```

### dotplot BP upregulated 2hr

```{r BP dotplot upregulated 2hr}
if(any(e_go_bp_2hr_up@result$p.adjust <0.01)){
  bp_dotplot_2hr_up <- dotplot(e_go_bp_2hr_up)+
    ggtitle("Dot Plot BP Upregulated 2hr")
  print(bp_dotplot_2hr_up)}
```

### cnet plot MF 2hr Upregulated

```{r mf cnet upregulated 2hr}
if(any(e_go_mf_2hr_up@result$p.adjust <0.01)){
  mf_cnetplot_2hr_up <- cnetplot(e_go_mf_2hr_up)+
    ggtitle("CNet Plot MF Upregulated 2hr")
  print(mf_cnetplot_2hr_up)}
```

### dotplot MF 2hr Upregulated

```{r mf dotplot upregulated 2hr}
if(any(e_go_mf_2hr_up@result$p.adjust <0.01)){
  mf_dotplot_2hr_up <- dotplot(e_go_mf_2hr_up)+
    ggtitle("Dot Plot MF Upregulated 2hr")
  print(mf_dotplot_2hr_up)}
```

### Cnet plot CC downregulated 2 hour
```{r cnet plot cc downregulated 2hr}
if(any(e_go_cc_2hr_down@result$p.adjust <0.01)){
  cc_cnet_2hr_down <- cnetplot(e_go_cc_2hr_down)+
    ggtitle("C Net Plot CC Down Regulated 2 hr")
  print(cc_cnet_2hr_down)}
```

### dotplot CC downregulated 2hr

```{r CC dotplot downregulated 2hr}
if(any(e_go_cc_2hr_down@result$p.adjust <0.01)){
  cc_dotplot_2hr_down <- dotplot(e_go_cc_2hr_down)+
    ggtitle("Dotplot CC Down regulated 2 hr")
  print(cc_dotplot_2hr_down)}
```

### cnet plot BP down regulated 2hr
```{r BP cnet down regulated 2hr}
if(any(e_go_bp_2hr_down@result$p.adjust <0.01)){
  bp_cnetplot_2hr_down <- cnetplot(e_go_bp_2hr_down)+
    ggtitle("CNet Plot BP down regulated 2hr")
  print(bp_cnetplot_2hr_down)}
```

### dotplot BP down regulated 2hr

```{r BP dotplot down regulated 2hr}
if(any(e_go_bp_2hr_down@result$p.adjust <0.01)){
  bp_dotplot_2hr_down <- dotplot(e_go_bp_2hr_down)+
    ggtitle("Dot Plot BP down regulated 2hr")
  print(bp_dotplot_2hr_down)}
```

### cnet plot MF 2hr down regulated

```{r mf cnet down regulated 2hr}
if(any(e_go_mf_2hr_down@result$p.adjust <0.01)){
  mf_cnetplot_2hr_down <- cnetplot(e_go_mf_2hr_down)+
    ggtitle("CNet Plot MF down regulated 2hr")
  print(mf_cnetplot_2hr_down)}
```

### dotplot MF 2hr down regulated

```{r mf dotplot down regulated 2hr}
if(any(e_go_mf_2hr_down@result$p.adjust <0.01)){
  mf_dotplot_2hr_down <- dotplot(e_go_mf_2hr_down)+
    ggtitle("Dot Plot MF down regulated 2hr")
  print(mf_dotplot_2hr_down)}
```

## GO 6hr

### GO 6hr setup
```{r clusterProfiler GO at 6hr,up and down}
e_go_cc_6hr_up <- cp_GO_analysis(up_6hr_ensembl, six_hr_dds, typeGO = "cc")

e_go_mf_6hr_up <- cp_GO_analysis(up_6hr_ensembl, six_hr_dds, typeGO = "mf")

e_go_bp_6hr_up <- cp_GO_analysis(up_6hr_ensembl, six_hr_dds, typeGO = "bp")

e_go_cc_6hr_down <- cp_GO_analysis(down_6hr_ensembl, six_hr_dds, typeGO = "cc")

e_go_mf_6hr_down <- cp_GO_analysis(down_6hr_ensembl, six_hr_dds, typeGO = "mf")

e_go_bp_6hr_down <- cp_GO_analysis(down_6hr_ensembl, six_hr_dds, typeGO = "bp")
```

### Cnet plot CC upregulated 6 hour
```{r cnet plot cc upregulated 6hr}
if(any(e_go_cc_6hr_up@result$p.adjust <0.01)){
  cc_cnet_6hr_up <- cnetplot(e_go_cc_6hr_up)+
    ggtitle("C Net Plot CC Upregulated 6 hr")
  print(cc_cnet_6hr_up)}
```

### dotplot CC upregulated 6hr

```{r CC dotplot upregulated 6hr}
if(any(e_go_cc_6hr_up@result$p.adjust <0.01)){
  cc_dotplot_6hr_up <- dotplot(e_go_cc_6hr_up)+
    ggtitle("Dotplot CC upregulated 6 hr")
  print(cc_dotplot_6hr_up)}
```

### cnet plot BP upregulated 6hr
```{r BP cnet upregulated 6hr}
if(any(e_go_bp_6hr_up@result$p.adjust <0.01)){
  bp_cnetplot_6hr_up <- cnetplot(e_go_bp_6hr_up)+
    ggtitle("CNet Plot BP Upregulated 6hr")
  print(bp_cnetplot_6hr_up)}
```

### dotplot BP upregulated 6hr

```{r BP dotplot upregulated 6hr}
if(any(e_go_bp_6hr_up@result$p.adjust <0.01)){
  bp_dotplot_6hr_up <- dotplot(e_go_bp_6hr_up)+
    ggtitle("Dot Plot BP Upregulated 6hr")
  print(bp_dotplot_6hr_up)}
```

### cnet plot MF 6hr Upregulated

```{r mf cnet upregulated 6hr}
if(any(e_go_mf_6hr_up@result$p.adjust <0.01)){
  mf_cnetplot_6hr_up <- cnetplot(e_go_mf_6hr_up)+
    ggtitle("CNet Plot MF Upregulated 6hr")
  print(mf_cnetplot_6hr_up)}
```

### dotplot MF 6hr Upregulated

```{r mf dotplot upregulated 6hr}
if(any(e_go_mf_6hr_up@result$p.adjust <0.01)){
  mf_dotplot_6hr_up <- dotplot(e_go_mf_6hr_up)+
    ggtitle("Dot Plot MF Upregulated 6hr")
  print(mf_dotplot_6hr_up)}
```

### Cnet plot CC downregulated 6 hour
```{r cnet plot cc downregulated 6hr}
if(any(e_go_cc_6hr_down@result$p.adjust <0.01)){
  cc_cnet_6hr_down <- cnetplot(e_go_cc_6hr_down)+
    ggtitle("C Net Plot CC Down Regulated 6 hr")
  print(cc_cnet_6hr_down)}
```

### dotplot CC downregulated 6hr

```{r CC dotplot downregulated 6hr}
if(any(e_go_cc_6hr_down@result$p.adjust <0.01)){
  cc_dotplot_6hr_down <- dotplot(e_go_cc_6hr_down)+
    ggtitle("Dotplot CC Down regulated 6 hr")
  print(cc_dotplot_6hr_down)}
```

### cnet plot BP down regulated 6hr
```{r BP cnet down regulated 6hr}
if(any(e_go_bp_6hr_down@result$p.adjust <0.01)){
  bp_cnetplot_6hr_down <- cnetplot(e_go_bp_6hr_down)+
    ggtitle("CNet Plot BP down regulated 6hr")
  print(bp_cnetplot_6hr_down)}
```

### dotplot BP down regulated 6hr

```{r BP dotplot down regulated 6hr}
if(any(e_go_bp_6hr_down@result$p.adjust <0.01)){
  bp_dotplot_6hr_down <- dotplot(e_go_bp_6hr_down)+
    ggtitle("Dot Plot BP down regulated 6hr")
  print(bp_dotplot_6hr_down)}
```

### cnet plot MF 6hr down regulated

```{r mf cnet down regulated 6hr}
if(any(e_go_mf_6hr_down@result$p.adjust <0.01)){
  mf_cnetplot_6hr_down <- cnetplot(e_go_mf_6hr_down)+
    ggtitle("CNet Plot MF down regulated 6hr")
 print(mf_cnetplot_6hr_down) }
```

### dotplot MF 6hr down regulated

```{r mf dotplot down regulated 6hr}
if(any(e_go_mf_6hr_down@result$p.adjust <0.01)){
  mf_dotplot_6hr_down <- dotplot(e_go_mf_6hr_down)+
    ggtitle("Dot Plot MF down regulated 6hr")
  print(mf_dotplot_6hr_down)}
```

