---
title: "YTH bound mRNAs w BDNF"
author: "Maddie Lombardo"
date: "2/15/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(PCAtools)
library(gprofiler2)
library(pcaExplorer)
library(clusterProfiler)
library(RColorBrewer)
library(ggpubr)
library(GOfuncR)
```


```{r read in files}
YTH_bound_mrna_file_path <- file.path(here::here("data","YTH bound mRNAs.csv"))

yth_bound_mrna <- read_csv(YTH_bound_mrna_file_path) %>% 
  janitor::clean_names()

new_rna_p_de_fp <- file.path(here::here("data","new_ratio_bayesian_p_de.csv"))

new_rna_p <- read_csv(new_rna_p_de_fp)
```

want to look at the genes downregulated with high fraction new mrna at 2hr to see if YTH is associated with the decay

lets see how many genes are bound to ythdf
Compare the N binding sites from POSTAR3 of these proteins to the log2Foldchange and mean_diff - probably a scatter plot
```{r genes bound to ythdf}
to_left_join <- yth_bound_mrna %>% 
  select(target_gene_id, binding_site_records, rbp, target_gene_symbol)

new_rna_p <- new_rna_p %>% 
  mutate(gene = gsub("\\..*", "", gene)) %>%
  left_join(to_left_join, by = c("gene" = "target_gene_id")) %>% 
  unique() %>% 
  mutate(binding_site_records = ifelse(is.na(binding_site_records),0,binding_site_records))
  
new_rna_p %>% ggplot(mapping = aes(log2FoldChange, binding_site_records))+
  geom_point()+
  facet_wrap(vars(time))

new_rna_p %>% 
  filter(binding_site_records > 5) %>% 
  ggplot(mapping = aes(mean_diff, binding_site_records))+
  geom_point()+
  facet_wrap(vars(time))+
  scale_y_continuous(trans = "log10")

new_rna_p %>%
  ggplot(mapping = aes(binding_site_records))+
                       geom_histogram()+
  scale_x_continuous(trans = "log10")
```


```{r filtering out the downregulated/high fraction new mrna at 2hr}
new_rna_p %>% 
  filter(time == 2) %>% 
  mutate(new_rna_sig = fct_relevel(new_rna_sig, "bdnf_lower_new_rna")) %>% 
  ggplot(mapping = aes(x = total_rna_sig, y = binding_site_records, fill = new_rna_sig))+
  geom_boxplot()+
  scale_y_continuous(trans = "log10")+
  ggpubr::stat_compare_means()

```

which genes have the most yth bound sites
overlap high_new_fraction and else with axonal vs soma 
```{r most yth}
yth_bound_mrna <- yth_bound_mrna %>% arrange(-binding_site_records)

axonal_genes_fp <- file.path(here::here("data","axonal_genes.txt"))

axonal_genes <- read_lines(axonal_genes_fp)
  
new_rna_p <- new_rna_p %>% 
  mutate(new_rna_lump = ifelse(new_rna_sig == "bdnf_higher_new_rna", "bdnf_higher_new_rna", "other")) %>%
  mutate(axonal_genes = target_gene_symbol %in% axonal_genes)

new_rna_p %>% 
  filter(time == 2) %>% 
  mutate(new_rna_sig = fct_relevel(new_rna_sig, "bdnf_lower_new_rna")) %>% 
  ggplot(mapping = aes(x = new_rna_lump, y = binding_site_records, fill = new_rna_lump))+
  geom_boxplot()+
  scale_y_continuous(trans = "log10")+
  ggpubr::stat_compare_means()+
  facet_wrap(vars(total_rna_sig))

new_rna_p %>% 
  ggplot(mapping = aes(x = mean_diff, color = axonal_genes))+
  stat_ecdf()+
  facet_wrap(vars(time))

new_rna_p %>% 
  ggplot(mapping = aes(x = log2FoldChange, color = axonal_genes))+
  stat_ecdf()+
  facet_wrap(vars(time))

new_rna_p %>% 
  ggplot(mapping = aes(x = mean_diff, color = axonal_genes))+
  stat_ecdf()+
  facet_grid(time ~ total_rna_sig)

```

